---
title: "Towards estimating true cholera burden: a systematic review and meta-analysis of Vibrio cholerae positivity"
author: "Kirsten E. Wiens, Hanmeng Xu, Kaiyue Zou, John Mwaba, Justin Lessler, Espoir B. Malembaka, Maya N. Demby, Godfrey Bwire, Firdausi Qadri, Elizabeth C. Lee, Andrew S. Azman"
output: 
  html_document:
    toc: true
    toc_float: true
---


<br>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      dev=c('png', 'pdf'), fig.path='figures/')
knitr::opts_knit$set(root.dir = here::here())
```

```{r preamble, warning = TRUE, message = FALSE}
# packages
pacman::p_load('tidyverse', 'data.table', 'readxl',
               'stringr', 'lubridate', 'reshape2',
               'boot', 'RVAideMemoire', 'meta', 
               'sf', 'raster', 'countrycode',
               'ggplot2', 'RColorBrewer', 'cowplot',
               'kableExtra', 'flextable')

# custom functions
source(here::here('code', 'utils.R'))

# JAGS settings
jags_run_id <- 'jags-cd'

# Stan settings
se_sp_draws <- 1000 # number of draws of sens/spec to use from the JAGS model output

# set random seed
set.seed(7341)
```

```{r load-clean-data, include = FALSE}
# shapefile
shp <- st_read(here::here('data/shapefiles/ne_50m_admin_0_countries/ne_50m_admin_0_countries.shp'))
shp <- shp %>% dplyr::select(ISO_A3, NAME_SORT, geometry)

#  load data
df_all <- read.csv(here::here('data', 'extracted_data', 'cc_extractions.csv'))
vars <- read.csv(here::here('data', 'extracted_data', 'extraction_variables.csv'))

# set column variable names
names(df_all) <- gsub('\\.', ' ', names(df_all))
df_all <- relocate(df_all, vars$Variable.label)
names(df_all) <- vars$Variable
df_all <- df_all[!is.na(names(df_all))]

# format columns
df_all <- df_all %>%
  # primary dataset
  filter(primary==1) %>%
  # dates
  mutate(outbreak_L = as.Date(outbreak_L, format = c('%d-%B-%Y')),
         outbreak_R = as.Date(outbreak_R, format = c('%d-%B-%Y')),
         TL = as.Date(TL, format = c('%d-%B-%Y')),
         TR = as.Date(TR, format = c('%d-%B-%Y'))) %>%
  # remove variable name text in parentheses and set factor levels
  mutate(study_design = gsub('\\s*\\([^\\)]+\\)', '', study_design),
         sampling_approach = gsub('\\s*\\([^\\)]+\\)', '', sampling_approach),
         study_design = factor(study_design, 
                               levels = c('Diagnostic test accuracy',
                                          'Vaccine effectiveness',
                                          'Surveillance', 'Other')),
         sampling_approach = factor(sampling_approach, 
                                    levels = c('All suspected cases', 'Systematic',
                                               'Simple random', 'Other', 
                                               'Convenience', 'Unknown')),
         sampling_quality = ifelse(sampling_approach %in% c('All suspected cases', 
                                                            'Systematic',
                                                            'Simple random'),
                                      'High', 'Low'),
         cases_dehydrated = ifelse(cases_dehydrated == 1, 'Yes', 'No')) %>%
  # region as defined by World Bank Development Indicators
  mutate(region = countrycode(country_iso3, 
                              origin = 'iso3c', 
                              destination = 'region')) %>%
  mutate(region = ifelse(region != 'Sub-Saharan Africa',
                         region, 
                         countrycode(country_iso3, 
                                     origin = 'iso3c', 
                                     destination = 'region23'))) %>%
  # surveillance is outbreak or non-outbreak
  mutate(surveillance_type = ifelse(surveillance_type == 'Outbreak surveillance',
                                    'Outbreak', 'Non-outbreak')) %>%
  # categorical version of age variable
  mutate(cases_min_age_categ = paste0('age', cases_min_age)) %>%
  # remove incomplete extractions
  filter(!is.na(sampling_approach))

# admin name editing (to match the shapefiles)
df_all <- edit_admin_names(df_all)
```

```{r match-polygons}
# match each entry_id to a polygon
shp_list <- readRDS(here::here("data", "shapefiles", "shp_list.RData"))
# shp_list contains three shapefiles (shp_admin0, shp_admin1, shp_admin2), e.g. if the minimal admin level of an entry is 0 (country level), the shapefile can be found in shp_admin0. If an entry contains more than one place, the were separated into different rows, each has its own polygon.
```

```{r load-incidence}
# load calculated estimated incidence (based on raster) from folder, the raster we have only covers sub-saharan Africa, other places were marked as NA
df_all <- suppressMessages(load_incidence(df_all%>%filter(country_iso3!='DZA'))) %>%
  # add Algeria data back with appropriate columns
  rbind(df_all %>% 
          filter(country_iso3=='DZA') %>% 
          mutate(admin_min_level=0, admin_min='Algeria', inc_calc=NA))
```

```{r aggregate-stratifications}
# variables to keep
keep_vars <- c('study_id', 'study_design', 'surveillance_type', 'case_definition', 'cases_min_age', 'cases_dehydrated', 'cases_hospitalized', 'sampling_quality', 'sampling_description', 'prop_lab_tested', 'num_tests', 'cholera_situation', 'diagnostic_test_type', 'diagnostic_test_name', 'region', 'country_iso3', 'admin0')

# aggregate stratifications by taking weighted mean
# this will be our main analysis dataset
df <- df_all %>%
  group_by_at(keep_vars) %>%
  summarize(prop_5 = weighted.mean(prop_5, n_cases_tested, na.rm=T),
            prop_15 = weighted.mean(prop_15, n_cases_tested, na.rm=T),
            prop_female = weighted.mean(prop_female, n_cases_tested, na.rm=T),
            prop_vaccinated = weighted.mean(prop_vaccinated, n_cases_tested, na.rm=T),
            prop_antibiotics = weighted.mean(prop_antibiotics, n_cases_tested, na.rm=T),
            prop_severe_dehyd = weighted.mean(prop_severe_dehyd, n_cases_tested, na.rm=T),
            prop_mod_dehyd = weighted.mean(prop_mod_dehyd, n_cases_tested, na.rm=T),
            n_cases_tested = sum(n_cases_tested, na.rm = TRUE),
            n_cases_positive = sum(n_cases_positive, na.rm = TRUE)) %>%
  ungroup()
```

<br>

## Study characteristics

### Figure S1. Data coverage by geography 

Number of study-periods in the primary dataset at each administrative level by country. An individual study has >1 "study-periods" if it reports cholera positivity for >1 time frame. Countries with >10 study-periods displayed as 10.

```{r figS1, fig.height = 10, fig.width = 7}
lt <- df_all %>% 
  # subset to primary dataset
  filter(primary==1) %>%
  # select relevant study-period columns
  dplyr::select(study_id, TL, TR, country_iso3, 
         admin0, admin1, admin2, admin3, place_name) %>%
  unique() %>%
  # only keep location names for lowest admin level
  mutate(admin0 = ifelse(admin1 == '' & admin2 == '' & admin3 == '', admin0, ''),
         admin1 = ifelse(admin2 == '' & admin3 == '', admin1, ''),
         admin2 = ifelse( admin3 == '', admin2, '')) %>%
  # reshape for counting
  reshape2::melt(measure.vars = c('admin0', 'admin1', 'admin2', 'admin3'),
                 id.vars = c('country_iso3')) %>%
  group_by(country_iso3, variable) %>%
  # count study-periods by admin level
  summarize(location_periods = sum(value!='')) %>%
  ungroup()

# plot for each admin level
figS1 <- list()
for (i in 0:3) {
  figS1[[i+1]] <- data_map(lt, shp, ad = paste0('admin', i))
}

cowplot::plot_grid(figS1[[1]], figS1[[2]], figS1[[3]], figS1[[4]],
                   labels = c('A', 'B', 'C', 'D'),
                   ncol = 1)


# # number at each level
# lt %>%
#   group_by(variable) %>%
#   # count study-periods by admin level
#   summarize(location_periods = sum(location_periods)) %>%
#   ungroup() %>%
#   pretty_table()

# # numbers for Haiti
# lt %>%
#   filter(country_iso3 == 'HTI') %>%
#   group_by(variable) %>%
#   # count study-periods by admin level
#   summarize(location_periods = sum(location_periods)) %>%
#   ungroup() %>%
#   pretty_table()
```

<br>

### Figure S2. Data coverage over time

Number of study-location-periods in the primary dataset at each administrative level by country within different time periods. An individual study has >1 "study-location-periods" if it reports cholera positivity for >1 time frame and/or for >1 administrative level. Countries with >10 study-location-periods displayed as 10. Year represents the year sampling was completed. Excludes 8 studies missing a study end date.

```{r figS2, fig.height = 10, fig.width = 7}
# similar to those above, but instead of stratification by admin level, stratify into 4 time periods (before 2000~2004, 2005~2009, 2010~2014, after 2015)
coverage_time <- df_all %>%
  filter(primary==1) %>%
  # select relevant study-period columns
  dplyr::select(study_id, TL, TR, country_iso3) %>%
  unique() %>%
  mutate(time_period = case_when(TR <= as.Date("2004-12-31") ~ "1997~2004",
                                 TR > as.Date("2004-12-31") & TR <= as.Date("2009-12-31") ~ "2005~2009",
                                 TR > as.Date("2009-12-31") & TR <= as.Date("2014-12-31") ~ "2010~2014",
                                 TR > as.Date("2014-12-31") ~ "2015~2022")) %>%
  # count by time (study ending time)
  count(country_iso3, time_period) %>% rename(variable = time_period, location_periods = n) %>%
  ungroup()

# plot for each time period
figS2 <- list()
for (i in c("1997~2004", "2005~2009", "2010~2014", "2015~2022")) {
  figS2[[i]] <- data_map(coverage_time, shp, ad = i)
}

cowplot::plot_grid(figS2[[1]], figS2[[2]], figS2[[3]], figS2[[4]],
                   labels = c('A', 'B', 'C', 'D'),
                   ncol = 1)

# # number at each level
# coverage_time %>%
#   group_by(variable) %>%
#   # count study-periods by admin level
#   summarize(location_periods = sum(location_periods)) %>%
#   ungroup() %>%
#   pretty_table()

# # numbers for Haiti
# coverage_time %>%
#   filter(country_iso3 == 'HTI') %>%
#   group_by(variable) %>%
#   # count study-periods by admin level
#   summarize(location_periods = sum(location_periods)) %>%
#   ungroup() %>%
#   pretty_table()
```

<br>

### Table 1. Study characteristics

Number of study-observations included in the dataset with each study characteristic. An individual study has >1 study-observations if it reports V. cholerae positivity for >1 sampling method or country.

```{r tab1}
# prep variables for table
sd <- df %>% 
  # format proportion tested
  mutate(prop_lab_tested = ifelse(prop_lab_tested < 0.05, '0-5',
                                   ifelse(prop_lab_tested >=0.05 & prop_lab_tested < 0.5, '5-50',
                                          ifelse(prop_lab_tested >= 0.5 & prop_lab_tested < 0.95, '50-95', '95+')))) %>%
  # number of tests
  mutate(num_tests = ifelse(num_tests >= 3, '3+', num_tests)) %>%
  # format sample sizes
  group_by(study_id, study_design, sampling_quality, prop_lab_tested, num_tests) %>%
  summarize(n = sum(n_cases_tested)) %>%
  ungroup() %>%
  mutate(n = ifelse(n < 10, '1-9', 
                    ifelse(n >=10 & n < 100, '10-99',
                           ifelse(n >= 100 & n < 1000, '100-999', '1000+')))) %>%
  unique() %>%
  dplyr::select(-study_id) 

# summarize and make table
sd %>%
  summarize_cats() %>%
  pretty_table()

sd %>%
  summarize_cats() %>%
  flextable() %>%
  save_as_docx(path = here::here('code/tables/table1.docx'))
```

<br>

### Table S2. Characteristics of suspected cholera cases reported by each study

```{r tabS2}
# summarize studies with and without covariates
cv <- df %>% 
  dplyr::select(study_id, prop_5, prop_15, prop_female, prop_mod_dehyd,
                prop_vaccinated, prop_antibiotics, prop_severe_dehyd) %>%
  mutate(prop_5_and_severe_dehyd = ifelse(!is.na(prop_5+prop_severe_dehyd), 'Yes', 'No'),
         prop_5_and_dehyd_and_antibiotics = ifelse(!is.na(prop_5+prop_severe_dehyd+prop_antibiotics), 'Yes', 'No')) %>%
  mutate(prop_under_5 = ifelse(!is.na(prop_5), 'Yes', 'No'),
         prop_under_15 = ifelse(!is.na(prop_15), 'Yes', 'No'),
         prop_female = ifelse(!is.na(prop_female), 'Yes', 'No'),
         prop_vaccinated = ifelse(!is.na(prop_vaccinated), 'Yes', 'No'),
         prop_antibiotics = ifelse(!is.na(prop_antibiotics), 'Yes', 'No'),
         prop_mod_dehyd = ifelse(!is.na(prop_mod_dehyd), 'Yes', 'No'),
         prop_severe_dehyd = ifelse(!is.na(prop_severe_dehyd), 'Yes', 'No')) %>%
  unique() %>%
  dplyr::select(-study_id, -prop_5, -prop_15)

# summarize and make table
cv %>%
  summarize_cats() %>%
  pretty_table()

cv %>%
  summarize_cats() %>%
  flextable() %>%
  save_as_docx(path = here::here('code/tables/tableS2.docx'))
```

<br>

## *V. cholerae* positivity in unadjusted data

### Figure 2. *Vibrio cholerae* positivity by study methodology and outbreak context

Proportion of suspected cholera cases that were confirmed positive by **A)** diagnostic test type, **B)** quality of sampling methods, where "high" includes all suspected cases or a random or stratified sample and "low" includes convenience or unreported sampling methods, **C)** age minimum in suspected case definition, where "0" indicates that no minimum age was set, and **D)** whether surveillance was initiated in response to an outbreak or whether it was routine surveillance or non-outbreak. Each point is a study-observation; individual studies may have multiple points if they reported positivity by multiple tests, reported results for multiple countries or outbreak contexts, and/or multiple sampling methods. Circles represent study-observations with high quality sampling and triangles indicate low quality sampling.

```{r}
cp <- df %>%
  # select cols
  group_by(study_id, study_design, sampling_quality,
           cases_min_age, cases_dehydrated, cases_hospitalized,
           surveillance_type, cholera_situation, 
           prop_antibiotics, prop_severe_dehyd,
           country_iso3, diagnostic_test_type) %>%
  # percent positive
  mutate(prop_pos = n_cases_positive/n_cases_tested,
         sampling_quality = ifelse(sampling_quality == 'High', 'High quality', 'Low quality')) %>%
  # hospitalized OR dehydrated
  mutate(cases_hosp_dehyd = ifelse(cases_hospitalized == 1 | cases_dehydrated == 1, 1, 0)) 
```

```{r}
# proportion positive by diagnostic test
f2a <- plot_posrate(cp, 'diagnostic_test_type', num_colors=3)

# cp %>% 
#   group_by(diagnostic_test_type) %>%
#   summarize(median = median(prop_pos)) %>%
#   ungroup %>%
#   pretty_table()

# proportion positive by sampling method
f2b <- plot_posrate(cp, 'sampling_quality', num_colors=3)

# cp %>% 
#   group_by(sampling_quality) %>%
#   summarize(median = median(prop_pos)) %>%
#   ungroup %>%
#   pretty_table()
```

```{r}
# proportion positive by case definitions
f2c <- plot_posrate(cp, 'cases_min_age', num_colors=5) + 
  ggtitle('Age minimum in case definition')

# cp %>% 
#   group_by(cases_min_age) %>%
#   summarize(median = median(prop_pos)) %>%
#   ungroup %>%
#   pretty_table()
```

```{r}
# proportion positive by surveillance type
f2d <- plot_posrate(cp, 'surveillance_type', num_colors=3)

# cp %>% 
#   group_by(surveillance_type) %>%
#   summarize(median = median(prop_pos)) %>%
#   ungroup %>%
#   pretty_table()
```

```{r fig2, fig.width = 8, fig.height = 6}
cowplot::plot_grid(f2a, f2b, f2c, f2d,
                   labels = c('A', 'B', 'C', 'D'),
                   nrow = 2)
```


<br>

### Figure S3. *Vibrio cholerae* positivity by incidence and suspected case characteristics

Relationship between reported *V. cholerae* positivity and **A)** proportion of suspected cholera cases tests that were under 5 years of age, **B)** suspected cholera incidence rate per 10,000 at study site, **C)** proportion of suspected cases severely dehydrated, and **D)** proportion of suspected cases on antibiotics prior to testing. Size of the points is proportional to the number of cases tested, and shapes indicate which diagnostic test was used to confirm *V. cholerae* infection. Confidence intervals for Spearman rank correlation coefficients estimated using bootstrapping (nrep=1000). Smoothing method is loess (without weights). In **B)** estimated suspected cholera incidence rates for 2010-2016 (Lessler et al., 2018) were aggregated to the administrative division that best represented each study’s catchment area by dividing the total estimated cholera cases in each area by its estimated population.

```{r}
# prep data
df_cont <- df %>%
  # select cols
  group_by(study_id, country_iso3, diagnostic_test_type,
           study_design, sampling_quality, surveillance_type,
           prop_antibiotics, prop_severe_dehyd,
           cases_min_age, prop_5) %>%
  # aggregate by study-period
  summarize(n_cases_tested = sum(n_cases_tested),
            n_cases_positive = sum(n_cases_positive)) %>%
  ungroup %>%
  # transform variables of interest
  mutate(posrate = n_cases_positive/n_cases_tested)
```

```{r}
# plot relationship between incidence and positivity rates
df_cor <- df_cont %>% 
  dplyr::select(study_id, posrate, sampling_quality, surveillance_type, 
                diagnostic_test_type, cases_min_age, n_cases_tested, prop_5) %>%
  na.omit()

ct <- spearman.ci(df_cor$posrate, df_cor$prop_5)

# plot relationship between walking time to health facilities and incidence
figS3a <- ggplot(df_cor, aes(y=posrate, x = prop_5)) +
  geom_point(alpha = 0.6, aes(size = n_cases_tested, shape = diagnostic_test_type)) +
  theme_bw() +
  theme(legend.position = 'bottom') +
  annotate('text', x = 0.5, y = 1.03, color = 'darkgreen',
           label = paste0('Spearman rho = ', signif(ct$estimate, 2), 
                          ' (95% Confidence Interval ', signif(ct$conf.int[[1]],2), 
                          ' to ', signif(ct$conf.int[[2]],2), ')')) +
  geom_smooth(method = 'loess',  color='purple')+
  labs(x = 'Proportion under 5', y = 'Proportion positive',
       size = NULL, shape = NULL)
```

```{r}
# plot relationship between incidence and positivity rates
df_cor <- df_all %>% 
  # select cols
  group_by(study_id, country_iso3, admin_min, inc_calc, diagnostic_test_type, 
           sampling_quality, surveillance_type, cases_min_age) %>%
  # aggregate
  summarize(n_cases_tested = sum(n_cases_tested),
            n_cases_positive = sum(n_cases_positive)) %>%
  ungroup %>%
  # transform variables of interest
  mutate(posrate = n_cases_positive/n_cases_tested,
         inc_calc_10000 = inc_calc * 10000) %>%
  na.omit()

ct <- spearman.ci(df_cor$posrate, df_cor$inc_calc_10000)

# plot relationship between positivity rate and incidence
figS3b <- ggplot(df_cor, aes(y=posrate, x = inc_calc_10000)) +
  geom_point(alpha = 0.6, aes(size = n_cases_tested, shape = diagnostic_test_type)) +
  scale_x_continuous(trans='log10') +
  theme_bw() +
  theme(legend.position = 'bottom') +
  annotate('text', x = 0.28, y = 1.02, color = 'darkgreen',
           label = paste0('Spearman rho = ', signif(ct$estimate, 2), 
                          ' (95% Confidence Interval ', signif(ct$conf.int[[1]],2), 
                          ' to ', signif(ct$conf.int[[2]],2), ')')) +
  geom_smooth(method = 'loess',  color='purple')+
  labs(x = 'Incidence (per 10,000)', y = 'Proportion positive',
       size = NULL, shape = NULL) 
```

```{r}
# plot relationship between incidence and positivity rates
df_cor <- df_cont %>% 
  dplyr::select(study_id, sampling_quality, surveillance_type, cases_min_age,
                diagnostic_test_type, n_cases_tested, posrate, prop_antibiotics) %>%
  na.omit()

ct <- spearman.ci(df_cor$posrate, df_cor$prop_antibiotics)

# plot relationship between positivity rate and incidence
figS3c <- ggplot(df_cor, aes(y=posrate, x = prop_antibiotics)) +
  geom_point(alpha = 0.6, aes(size = n_cases_tested, shape = diagnostic_test_type)) +
  theme_bw() +
  theme(legend.position = 'bottom') +
  scale_size_continuous(breaks = c(1000, 3000, 5000)) +
  annotate('text', x = 0.23, y = 1.02, color = 'darkgreen',
           label = paste0('Spearman rho = ', signif(ct$estimate, 2), 
                          ' (95% Confidence Interval ', signif(ct$conf.int[[1]],2), 
                          ' to ', signif(ct$conf.int[[2]],2), ')')) +
  geom_smooth(method = 'loess',  color='purple')+
  labs(x = 'Proportion on antibiotics', y = 'Proportion positive',
       size = NULL, shape = NULL) 
```

```{r}
# plot relationship between incidence and positivity rates
df_cor <- df_cont %>% 
  dplyr::select(study_id, sampling_quality, surveillance_type, cases_min_age,
                diagnostic_test_type, n_cases_tested, posrate, prop_severe_dehyd) %>%
  na.omit()

ct <- spearman.ci(df_cor$posrate, df_cor$prop_severe_dehyd)

# plot relationship between positivity rate and incidence
figS3d <- ggplot(df_cor, aes(y=posrate, x = prop_severe_dehyd)) +
  geom_point(alpha = 0.6, aes(size = n_cases_tested, shape = diagnostic_test_type)) +
  theme_bw() +
  theme(legend.position = 'bottom') +
  scale_size_continuous(breaks = c(100, 300, 500)) +
  annotate('text', x = 0.45, y = 1.02, color = 'darkgreen',
           label = paste0('Spearman rho = ', signif(ct$estimate, 2), 
                          ' (95% Confidence Interval ', signif(ct$conf.int[[1]],2), 
                          ' to ', signif(ct$conf.int[[2]],2), ')')) +
  geom_smooth(method = 'loess',  color='purple')+
  labs(x = 'Proportion severely dehydrated', y = 'Proportion positive',
       size = NULL, shape = NULL) 
```

```{r figS3, fig.width = 10, fig.height = 10}
cowplot::plot_grid(figS3a, figS3b, figS3d, figS3c,
                   nrow = 2,
                   labels = c('A', 'B', 'C', 'D'))
```

<br>

## Adjusted underlying *V. cholerae* positivity

```{r jags-validation-data}
# Re-create dataset using validation data from Fig 2, Table 4, and text of Debes et al. 2021
# https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8176501/
df_ken <- rbind(
  data.frame(id = 1:8, culture=1, pcr=0, rdt=1),
  data.frame(id = 1:69, culture=1, pcr=1, rdt=1),
  data.frame(id = 1:2, culture=0, pcr=1, rdt=1),
  data.frame(id = 1:2, culture=0, pcr=1, rdt=0),
  data.frame(id = 1:149, culture=0, pcr=0, rdt=0)
)
df_ken$id <- NULL

# Re-create dataset using validation data from Table 2 and Fig 3 of Sayeed et al. 2018
# The results for Cholkit and Crystal VC are the same comapred to the other tests
# https://journals.plos.org/plosntds/article?id=10.1371/journal.pntd.0006286
df_bgd <- rbind(
  data.frame(id = 1:15, culture=1, pcr=1, rdt=1),
  data.frame(id = 1:6, culture=0, pcr=1, rdt=1),
  data.frame(id = 1:4, culture=1, pcr=0, rdt=1),
  data.frame(id = 1:5, culture=0, pcr=0, rdt=1),
  data.frame(id = 1:47, culture=0, pcr=0, rdt=0)
)
df_bgd$id <- NULL

# Validation data from Table 1 of Ontweka et al. 2016 PLOS ONE
# https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0168257
df_ssd <- read_xlsx(here::here('data', 'raw_data', 'pone.0168257.s002.xlsx')) %>%
  dplyr::select(`...12`, `...14`, `...9`) 

names(df_ssd) <- c('culture', 'pcr', 'rdt')
df_ssd <- df_ssd[-c(1,2),]

df_ssd <- df_ssd %>%
  mutate(culture = ifelse(culture=='Negative', 0, 1),
         rdt = as.numeric(rdt),
         rdt = ifelse(rdt==2, 1, rdt),
         pcr = as.numeric(pcr)) %>%
  na.omit()

# Validation data from Mwaba et al. 2018
# https://onlinelibrary.wiley.com/doi/full/10.1111/tmi.13084
df_zmb_pcr <- read_xlsx(here::here('data', 'raw_data', 'Lusaka_2016_PCR.xlsx')) %>%
 dplyr::select(Sample, `Final result`) %>%
 na.omit() %>%
 mutate(`Final result` = ifelse(`Final result` == '+', 1, 0))

df_zmb_rdt <- read_xlsx(here::here('data', 'raw_data', 'Lusaka_2016_RDT.xlsx'),
                       range = 'C4:L237') %>%
 # Included in the paper
 dplyr::filter(Include==1) %>%
 mutate(Sample = as.numeric(gsub('o', '', ID))) %>%
 mutate(`Direct RDT` = ifelse(`APW RDT` == '+' | `Direct RDT` == 'Pos', 1, 0),
        Culture = ifelse(Culture == '+' | Culture == 'Pos', 1, 0))

df_zmb <- merge(df_zmb_rdt%>%dplyr::select(Sample, Culture, `Direct RDT`),
               df_zmb_pcr) %>%
 dplyr::select(Culture, `Final result`, `Direct RDT`)

names(df_zmb) <- c('culture', 'pcr', 'rdt')
rm(df_zmb_pcr, df_zmb_rdt)

# create parent dataset and vector with study label
df_val <- rbind(df_ken, df_bgd, 
                df_ssd, df_zmb
                )
val_id <- c(rep(1, nrow(df_ken)),
            rep(2, nrow(df_bgd)),
            rep(3, nrow(df_ssd)),
            rep(4, nrow(df_zmb))
            )
```

```{r jags-model-results}
# file to save to
jags_est_file <- here::here('data', 'generated_data', 'sens_spec_estimates', 
                            paste0('sens-spec-', jags_run_id, '.rds'))
jags_fit <- readRDS(jags_est_file)

# randomly select draws to run in stan
draws <- round(runif(se_sp_draws, 1, 4000))
se_draws <- jags_fit$BUGSoutput$sims.list$mu_se[draws, ]
sp_draws <- jags_fit$BUGSoutput$sims.list$mu_sp[draws, ]
colnames(se_draws) <- colnames(sp_draws) <- c('Culture', 'PCR', 'RDT')

# also save study-specific estimates
# first index is draw, second is test, third is study
se_study <- jags_fit$BUGSoutput$sims.list$se[draws, , ]
sp_study <- jags_fit$BUGSoutput$sims.list$sp[draws, , ]
```

### Table S3. Estimated sensitivity and specificity of each diagnostic test

Estimate is median sensitivity and specificity pooled across four studies that reported results for all diagnostics tests, as described in Methods. Parentheses show 95% Credible Interval.

```{r tabS3}
# extract sensitivity and specificity estimates
se_sp_est <- jags_fit$BUGSoutput$summary[2:7, c('2.5%', '50%', '97.5%')]
se_sp_est <- round(se_sp_est, 3)*100

# table S2
tabS3 <- data.frame('Measure' = c(rep('Sensitivity', 3), rep('Specificity', 3)),
                    'Test' = rep(c('Culture', 'PCR', 'RDT'), 2),
                    'Estimate' = paste(se_sp_est[,2],'(',se_sp_est[,1],'-',se_sp_est[,3],')'))

names(tabS3) <- c('Measure', 'Test', 'Estimate (%)')
pretty_table(tabS3)

tabS3 %>%
  flextable() %>%
  save_as_docx(path = here::here('code/tables/tableS3.docx'))
```

```{r epi-data-stan}
# set covariates
covs <- c('sampling_quality', 'cases_min_age', 'cases_dehydrated', # adjust for these
          'surveillance_type') # examine differences by these (and age categories in separate analysis)

# positivity rates from meta-analysis primary dataset
sr_dat <- df %>%
  group_by_at(c('study_id', 'country_iso3', 'region', covs,
                'diagnostic_test_type', 'diagnostic_test_name')) %>%
  summarize(n_cases_tested = sum(n_cases_tested),
            n_cases_positive = sum(n_cases_positive)) %>%
  ungroup()

# cast number tested by test type
test_dat <- sr_dat %>%
  reshape2::dcast(as.formula(paste0(paste0(c('study_id', 'country_iso3', 'region', covs), 
                                           collapse = '+'), 
                                    '~ diagnostic_test_type')),
                  value.var = 'n_cases_tested', fill = 0)

# cast number positive by test type
pos_dat <- sr_dat %>%
  reshape2::dcast(as.formula(paste0(paste0(c('study_id', 'country_iso3', 'region', covs), 
                                           collapse = '+'), 
                                    '~ diagnostic_test_type')),
                  value.var = 'n_cases_positive', fill = 0)

# add random effect label by observation
test_dat <- test_dat %>% mutate(study_lab = 1:nrow(test_dat))
pos_dat <- pos_dat %>% mutate(study_lab = 1:nrow(test_dat))
```

```{r}
proppos <- data.frame(run_id = NA,
                      stan_id = NA,
                      median = NA,
                      lower = NA,
                      upper = NA,
                      row.names = NULL)

file_names <- list.files(here::here('data', 'generated_data', 'adjusted_positivity_rates'))

for (i in file_names) {
  
  posrate <- readRDS(here::here('data', 'generated_data', 'adjusted_positivity_rates', i))
  model_name <- gsub('posrate-|-full|-d1000|\\.rds', '', i)
  
  # save positivity rates results
  proppos <- rbind(proppos,
                   data.frame(run_id = model_name,
                              stan_id = 'proportion positive',
                              median = median(posrate$mu_p, na.rm = T),
                              lower = quantile(posrate$mu_p, probs = .025, na.rm = T),
                              upper = quantile(posrate$mu_p, probs = .975, na.rm = T),
                              row.names = NULL
                              )
                   )
  
  # save beta coefficients
  for (c in names(posrate$beta)) {
    if (c != '(Intercept)') {
      proppos <- rbind(proppos,
                       data.frame(
                         run_id = model_name,
                         stan_id = c,
                         median = median(posrate[['beta']][[c]], na.rm = T),
                         lower = quantile(posrate[['beta']][[c]], probs = .025, na.rm = T),
                         upper = quantile(posrate[['beta']][[c]], probs = .975, na.rm = T),
                         row.names = NULL
                       )
      )
    }
  }
  
}

# tidy results
proppos <- proppos %>%
  drop_na() %>%
  mutate(median=round(median,2),
         lower=round(lower,2),
         upper=round(upper,2))
```

<br>

### Figure S4. Posterior distributions of *V. cholerae* positivity

Posterior distributions of underlying true *V. cholerae* positivity estimated using the random-effects model adjusted for diagnostic test sensitivity and specificity and study methods (i.e., sampling quality and age minimum in case definition), corresponding to results in Table 2. **A)** Prior (green) and posterior distribution (purple) of positivity in logit space. **B)** Prior (green) and posterior predictive distribution (purple) of positivity in logit space. **C)** Histogram of estimated positivity, or the true underlying proportion positive, based on studies in the meta-analysis. **D)** Histogram of the posterior predictive distribution of positivity, or what we predict the underlying proportion positive would be in a new hypothetical study site.

```{r figS4, fig.width = 8, fig.height=6}
# distribution of percent positive
posrate <- readRDS(here::here('data', 'generated_data', 'adjusted_positivity_rates',
                              paste0('posrate-mod2-d1000-full.rds')))

s4a <- plot_prior_post(logit(posrate$mu_p), 
                'Proportion positive, adjusted for methods', 
                'logit(proportion positive)',
                arg1=0, arg2=1, dprior='dnorm')

s4b <- plot_prior_post(logit(posrate$p_pred$p_pred), 
                'Posterior predictive distribution', 
                'logit(proportion positive)',
                arg1=0, arg2=1, dprior='dnorm')

# get quantiles
mu_quant <- quantile(posrate$mu_p, c(0.025, 0.5, 0.975))

s4c <- ggplot() +
  geom_histogram(aes(posrate$mu_p))+
  xlim(c(0,1)) +
  theme_classic() +
  xlab('Proportion positive')

pred_quant <- quantile(posrate$p_pred$p_pred, c(0.025, 0.5, 0.975))

s4d <- ggplot() +
  geom_histogram(aes(posrate$p_pred$p_pred)) +
  xlim(c(0,1)) +
  theme_classic() + 
  theme_classic() + xlab('Posterior predictive distribution')

# plot
cowplot::plot_grid(s4a, s4b, s4c, s4d, labels = c('A', 'B', 'C', 'D'))

# save
message(pred_quant)
prop_pos_pred <- posrate$p_pred$p_pred
save(prop_pos_pred, file = here::here('data', 'generated_data', 'predictions', 'prop_pos_pred.RData'))
```

<br>

### Table S4. Estimated underlying *V. cholerae O1/O139* positivity

"Unadjusted" is mean percent positive (95% confidence interval) from random effects meta-analysis implemented using R package "meta" without any adjustments. "Adjusted" is the underlying true percent positive (95% credible interval) estimated using the Bayesian hierarchical model, adjusted for sensitivity/specificity of the tests, age in case definition, sampling quality, and whether or not surveillance was initiated in response to an outbreak. Also includes sensitivity analysis of increasing variance on all priors

```{r}
df_meta <- df %>%
  group_by(study_id, country_iso3, study_design,
           sampling_quality, diagnostic_test_type, diagnostic_test_name) %>%
  summarize(n_cases_tested = sum(n_cases_tested),
            n_cases_positive = sum(n_cases_positive)) %>%
  ungroup() %>%
  mutate(study_lab = paste0(country_iso3, '-', study_id)) %>%
  arrange(country_iso3, study_id)

ma_unadj <- metaprop(n_cases_positive, n_cases_tested,
                     studlab = df_meta$study_lab, data = df_meta)
```

```{r tabS4}
# results from RE meta-analysis in metafor with no adjustments
pr1 <- c(0.37, 0.41, 0.46)

# median adjusted proportion positive
pr2 <- proppos %>%
  filter(run_id %in% c('mod1', 'mod2', 'mod3', 'mod3-sa'), stan_id == 'proportion positive') %>%
  dplyr::select(lower, median, upper) %>%
  round(2)

pr2 <- rbind(pr1, pr2)
pr2 <- pr2*100

pr2 <- pr2 %>%
  mutate(variable = c('Unadjusted', 
                      'Adjusted:\ntests',
                      'Adjusted:\ntests,\nmethods',
                      'Adjusted:\ntests,\nmethods,\noutbreak',
                      'Adjusted:\ntests,\nmethods,\noutbreak,\ntest priors')) %>%
  mutate(variable = factor(variable, levels = c('Unadjusted', 
                      'Adjusted:\ntests',
                      'Adjusted:\ntests,\nmethods',
                      'Adjusted:\ntests,\nmethods,\noutbreak',
                      'Adjusted:\ntests,\nmethods,\noutbreak,\ntest priors')))

# table S4
tabS4 <- data.frame('Version' = c('Unadjusted', 
                                 'Adjusted for test performance',
                                 'Adjusted for test performance & methods',
                                 'Adjusted for test performance, methods, & outbreak setting',
                                 'Adjusted for test performance, methods, & outbreak setting, increased variance on priors'),
                   'Positivity rate' = paste(pr2[,2],'(',pr2[,1],'-',pr2[,3],')'))

names(tabS4) <- c('Version', 'Positivity (%)')
pretty_table(tabS4)

tabS4 %>%
  flextable() %>%
  save_as_docx(path = here::here('code/tables/tableS4.docx'))
```

<br>

### Figure 3. Estimated underlying *V. cholerae* positivity

**A)** Posterior distributions of pooled percent sensitivity and specificity of culture (top), PCR (middle), and RDT (bottom) for detecting V. cholerae O1O139 infections in suspected cholera cases. Dashed lines represent median values of each distribution. **B)** "Unadjusted" is mean percent positive (95% confidence interval) from random effects meta-analysis implemented using R package "meta" without any adjustments. "Adjusted" is the underlying true percent positive (95% credible interval) estimated using the Bayesian hierarchical model, adjusted for sensitivity/specificity of the tests, age in case definition, sampling quality, and whether or not surveillance was initiated in response to an outbreak.

```{r fig3, fig.width = 6.5, fig.height = 4.5}
# set colors
pal <- c('grey30', brewer.pal(n=4, name='Dark2'))

# positivity plot
f3b <- ggplot(pr2) +
          geom_point(aes(x = variable, y = median, color = variable), size = 2) +
          geom_errorbar(aes(x = variable, ymin = lower, ymax = upper, color = variable), width = 0.3) +
          scale_color_manual(values = pal) +
          theme_bw() +
          theme(legend.position = 'none') +
          ylab('Positivity (%)') + xlab(NULL) +
          ggtitle('Unadjusted and adjusted positivity')

# posterior of sensitivity and specificity
pal <- brewer.pal(11, 'PRGn')

se_long <- melt(se_draws) %>%
  mutate(measure = 'Sensitivity')
sp_long <- melt(sp_draws) %>%
  mutate(measure = 'Specificity')

long <- rbind(se_long, sp_long)

# medians
ss_meds <- long %>% 
  group_by(Var2, measure) %>%
  summarize(value = median(value)*100) %>%
  ungroup()

f3a <- ggplot(long) + 
  geom_vline(data=ss_meds%>%filter(measure == 'Sensitivity'), aes(xintercept = value), 
             color = pal[9], linetype = 'dashed') +
  geom_vline(data=ss_meds%>%filter(measure == 'Specificity'), aes(xintercept = value), 
             color = pal[2], linetype = 'dashed') +
  geom_density(aes(value*100, color = measure)) + 
  theme_classic() +
  facet_wrap('Var2', ncol = 1) +
  xlab(NULL) + ylab('Posterior Density') + 
  scale_color_manual(values = c(pal[9], pal[2]),
                     name = NULL, 
                     labels = c('Sensitivity', 'Specificity')) +
  theme(legend.position = 'bottom',
        legend.background=element_blank())

cowplot::plot_grid(f3a, f3b,
                   labels = c('A', 'B'),
                   ncol = 2, rel_widths = c(0.7, 1))
```

<br>

### Figure 4. Forest plot of study estimates and underlying positivity

Black points indicate median study-level underlying positivity and 95% Credible Interval (CI). Teal, orange, and purple points indicate the proportion positive reported by study for culture, PCR, and RDT, respectively, and corresponding error bars indicate 95% confidence interval for a binomial probability. Studies are labeled by country ISO3 code, whether or not they used high quality sampling methods, and whether a minimum age was set in the suspected cholera case definition. Green distribution indicates median and 95% CI of pooled positivity. Pink distribution indicates median and 95% CI of posterior predictive distribution of positivity. Studies are split into outbreak and non-outbreak for ease of interpretation, but the green and pink distributions represent the global estimates across all studies. 

```{r}
# underlying positivity rate by study
df_p <- posrate$p 

# get median and 95% CI by study
study_est <- data.frame(matrix(nrow=0, ncol=9))
names(study_est) <- c('study_id', 'country_iso3', 'sampling_quality', 'cases_min_age',
                      'surveillance_type', 'median', 'lower', 'upper')
  
for (i in 1:ncol(df_p)) {
  study_est <- rbind(study_est,
                     data.frame(study_id = test_dat[i, 'study_id'],
                                country_iso3 = test_dat[i, 'country_iso3'],
                                sampling_quality = test_dat[i, 'sampling_quality'],
                                cases_min_age = test_dat[i, 'cases_min_age'],
                                surveillance_type = test_dat[i, 'surveillance_type'],
                                median = median(df_p[, i][[1]], na.rm = T),
                                lower = quantile(df_p[, i][[1]], probs = .025, na.rm = T),
                                upper = quantile(df_p[, i][[1]], probs = .975, na.rm = T)))
}

# add original study data
study_est <- left_join(study_est, test_dat) %>%
  dplyr::select(-study_lab) %>%
  # proportion positive by test
  mutate(Culture_p = pos_dat$Culture/test_dat$Culture,
         PCR_p = pos_dat$PCR/test_dat$PCR,
         RDT_p = pos_dat$RDT/test_dat$RDT) %>%
  group_by(study_id, country_iso3, sampling_quality, cases_min_age, surveillance_type,
           median, upper, lower, Culture_p, PCR_p, RDT_p) %>%
  # binomial confidence intervals and variance for proportions
  summarize(Culture_lower = binomial_ci(Culture_p, Culture)[[1]],
            Culture_upper = binomial_ci(Culture_p, Culture)[[2]],
            PCR_lower = binomial_ci(PCR_p, PCR)[[1]],
            PCR_upper = binomial_ci(PCR_p, PCR)[[2]],
            RDT_lower = binomial_ci(RDT_p, RDT)[[1]],
            RDT_upper = binomial_ci(RDT_p, RDT)[[2]],
            Culture_v = prop_var(Culture_p, Culture),
            PCR_v = prop_var(PCR_p, PCR),
            RDT_v= prop_var(RDT_p, RDT),
            # maximum variance across studies that use multiple tests
            #TODO check that this is OK
            vi = max(c(Culture_v, PCR_v, RDT_v), na.rm=T)) %>%
  ungroup()
```

```{r i2, message = TRUE}
# calculate tau2, between-study heterogeneity
# https://www.frontiersin.org/articles/10.3389/fvets.2020.00271/full
# https://journals.sagepub.com/doi/full/10.1177/25152459211031256
# taken from the Stan printed output
tau2 <- inv.logit(1.3)

# grab v, within-study sampling variance, dropping the study that reported 0 positives
vi <- study_est%>%filter(vi!=0)%>%dplyr::select(vi)

# calculate i2
# https://www.metafor-project.org/doku.php/tips:i2_multilevel_multivariate#:~:text=For%20a%20standard%20random%2Deffects,ith%20i%20t
k <- nrow(test_dat)-1 # dropped 1 study with 0 positives
sum_wi <- sum(1/vi[[1]])
sum_wi2 <- sum((1/vi[[1]])^2)
v <- ((k-1)*sum_wi)/(sum_wi^2-sum_wi2)
i2 <- 100*tau2/(tau2+v)

message('I2: ', round(i2, 2), '; tau2: ', round(tau2, 2))
```

```{r fig4, fig.height = 8.5, fig.width = 8.5}
# plot from lowest to highest median positivity
setorderv(study_est, 'median')

# create study labels
study_est <- study_est %>%
  mutate(study_lab = paste0(country_iso3, ' - ',
                            ifelse(sampling_quality=='High', 'High - ', 'Low  - '),
                            ifelse(cases_min_age==10, paste0('', cases_min_age),
                                   paste0(' ', cases_min_age, ' '))))

# forest plot for outbreak surveillance
out <- forest_plot(study_est %>% 
                     filter(surveillance_type=='Outbreak') %>%
                     mutate(id = 1:nrow(study_est %>% filter(surveillance_type=='Outbreak')))) +
  ggtitle('Outbreak surveillance')

# forest plot for non-outbreak surveillance
nonout <- forest_plot(study_est %>% 
                        filter(surveillance_type=='Non-outbreak') %>%
                        mutate(id = 1:nrow(study_est %>% filter(surveillance_type=='Non-outbreak')))) +
  ggtitle('Non-outbreak surveillance')

# plot together
cowplot::plot_grid(out, nonout) 
```

<br>

## Factors associated with variation in *V. cholerae* positivity

### Table S5. Odds of *V. cholerae* positivity by age and outbreak context

Odds that a suspected cholera case seeking testing or care has a true _V. cholerae_ O1/O139 infection with low sampling quality compared to high sampling quality, with increasing minimum age set in suspected case definition, and with surveillance initiated in response to an outbreak compared to non-outbreak surveillance  (i.e., routine or post-vaccination surveillance). 

```{r tabS5}
# median coefficients
c1 <- proppos %>%
  filter(run_id %in% c('mod3'),
         stan_id == 'sampling_qualityLow' | stan_id == 'cases_min_age' | stan_id == 'surveillance_typeOutbreak') %>%
  dplyr::select(lower, median, upper) %>%
  round(2) %>%
  mutate(variable = c('Low sampling quality',
                      'Minimum age in\ncase definition',
                      'Outbreak surveillance'))

# table S5
tabS5 <- data.frame('Variable' = c('Low sampling quality',
                                  'Minimum age in case definition',
                                  'Outbreak surveillance'),
                   'Odds ratio' = paste(c1[,2],'(',c1[,1],'-',c1[,3],')'))

names(tabS5) <- c('Variable', 'Odds ratio')
pretty_table(tabS5)

tabS5 %>%
  flextable() %>%
  save_as_docx(path = here::here('code/tables/tableS5.docx'))
```

<br>

### Table S6. Odds of *V. cholerae* positivity by categorical age in case definition

Odds that a suspected cholera case seeking care has a true _V. cholerae_ infection by minimum age set in suspected case defintion. Coefficients correspond to random effects models run with age as a categorical variable (reference group = no age minimum age explicitly set in case definition or 0). All other model settings and covariates are the same as in Table 3.

```{r tabS6}
# median coefficients
c2 <- proppos %>%
  filter(run_id %in% c('mod3c'), 
         stan_id %like% 'cases_min_age' | stan_id %in% c('sampling_qualityLow', 'surveillance_typeOutbreak')) %>%
  mutate(stan_id = gsub('cases_min_age_categage', '', stan_id),
         stan_id = gsub('sampling_qualityLow', '0.1', stan_id),
         stan_id = gsub('surveillance_typeOutbreak', '100', stan_id),
         stan_id = as.numeric(stan_id)) %>%
  arrange(stan_id) %>%
  dplyr::select(lower, median, upper) %>%
  round(2)

# table S6
tabS6 <- data.frame('Variable' = c('Low sampling quality', 
                                   paste('Min age in case definition:', c('1','2','5','10')),
                                   'Outbreak surveillance'),
                    'Odds ratio' = paste(c2[,2],'(',c2[,1],'-',c2[,3],')'))

names(tabS6) <- c('Variable', 'Odds ratio')
pretty_table(tabS6)

tabS6 %>%
  flextable() %>%
  save_as_docx(path = here::here('code/tables/tableS6.docx'))
```

<br>

### Table S7. Sensitivity analysis of the odds of *V. cholerae* positivity by age and outbreak context

Odds that a suspected cholera case seeking testing or care has a true _V. cholerae_ O1/O139 infection with low sampling quality compared to high sampling quality, with increasing minimum age set in suspected case definition, and with surveillance initiated in response to an outbreak compared to non-outbreak surveillance  (i.e., routine or post-vaccination surveillance) in models where variance was increased on all priors. 

```{r tabS7, eval = TRUE}
# median coefficients
c1s <- proppos %>%
  filter(run_id %in% c('mod3-sa'),
         stan_id == 'sampling_qualityLow' | stan_id == 'cases_min_age' | stan_id == 'surveillance_typeOutbreak') %>%
  dplyr::select(lower, median, upper) %>%
  round(2)

# table S7
tabS7 <- data.frame('Variable' = c('Low sampling quality',
                                   'Minimum age in case definition',
                                   'Outbreak surveillance'),
                   'Odds ratio' = paste(c1[,2],'(',c1[,1],'-',c1[,3],')'))

names(tabS7) <- c('Variable', 'Odds ratio')
pretty_table(tabS7)

tabS7 %>%
  flextable() %>%
  save_as_docx(path = here::here('code/tables/tableS7.docx'))
```
